\documentclass[12pt]{report}
\usepackage{xcolor}

\title{\textbf{ByteRoute} \\ \large ASW Report}

\author{
  Babini Stefano\\
  0001125127 \\
  \texttt{stefano.babini5@studio.unibo.it}
}

\date{\today}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage[english]{babel}
\usepackage{float}
\pagestyle{plain}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage[export]{adjustbox}
\usetikzlibrary{petri,positioning,shapes.geometric,arrows,fit,backgrounds}
\usepackage{listings}
\usepackage{titlesec}
\usepackage{booktabs}
\usepackage{longtable}

\titleformat{\chapter}[display]{\normalfont\bfseries}{}{0pt}{\Large}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    filecolor=magenta,
    urlcolor=cyan,
}

\graphicspath{{images/}}

\renewcommand{\lstlistingname}{Code}
\renewcommand{\lstlistlistingname}{List of \lstlistingname s}

% Color setup
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

% Code listing style
\lstdefinestyle{mystyle}{
  backgroundcolor=\color{backcolour},
  commentstyle=\color{codegreen},
  keywordstyle=\color{magenta},
  numberstyle=\tiny\color{codegray},
  stringstyle=\color{codepurple},
  basicstyle=\ttfamily\footnotesize,
  breakatwhitespace=false,
  breaklines=true,
  captionpos=b,
  keepspaces=true,
  numbers=left,
  numbersep=5pt,
  showspaces=false,
  showstringspaces=false,
  showtabs=false,
  tabsize=2
}

\lstset{style=mystyle}

\begin{document}

\maketitle
\newpage
\tableofcontents{}
\newpage

% =========================================================
\chapter{Introduction}
% =========================================================

ByteRoute is a real-time network traffic monitoring platform developed as a university project
for the ASW (\emph{Applicazioni e Servizi Web}) course. Its core objective is to give users
live, geolocated visibility into the internet destinations their computer reaches: which servers,
autonomous systems, and countries receive their traffic, and in what volume.

The system is built as a pnpm monorepo containing three independently deployable components:
a lightweight Linux packet capture agent written in Go, a Node.js/Express backend with
GeoIP enrichment and Socket.IO broadcasting, and a Vue.js~3 single-page dashboard. All
components are packaged as Docker images and integrated through Docker Compose behind a
Traefik reverse proxy, providing a production grade topology that can be started with a
single command.

The remainder of this report is organised as follows.
\begin{itemize}
      \item Chapter~\ref{ch:requirements} describes user, functional, and non-functional requirements.
      \item Chapter~\ref{ch:design} presents the overall system architecture, data flow, domain model, and
              design decisions.
      \item Chapter~\ref{ch:technologies} surveys the key technologies employed.
      \item Chapter~\ref{ch:implementation} details notable implementation aspects of each component.
      \item Chapter~\ref{ch:testing} discusses testing strategy and coverage results.
      \item Chapter~\ref{ch:deployment} describes how the system is deployed.
      \item Chapter~\ref{ch:conclusion} draws conclusions and outlines future work.
\end{itemize}

% =========================================================
\chapter{Requirements}
\label{ch:requirements}
% =========================================================

\section{User Requirements}

The target users of ByteRoute are technically literate individuals (developers, system
administrators, security researchers) who want transparent, real-time insight into the network
activity of their workstations or servers.

\begin{itemize}
    \item Register an account on the platform and manage personal credentials.
    \item Create one or more \emph{tenants} (logical monitoring contexts, e.g.\ one per host).
    \item View a live world map showing where outbound connections are directed.
    \item Inspect an up to date list of active connections enriched with country, city, and
          ASN information.
    \item Browse statistical breakdowns of traffic by country, ASN, and protocol.
    \item Examine time series charts of connection volume.
    \item Generate client tokens so that the Go agent can post data without
          exposing the user's main JWT.
    \item Switch between tenants to inspect different monitored hosts from the same dashboard.
\end{itemize}

\newpage
\section{Functional Requirements}

\begin{enumerate}
    \item \textbf{Packet capture}: The agent must capture IP packets on a specified network
          interface using libpcap and derive $5$-tuple connection records.
    \item \textbf{Deduplication}: The agent must deduplicate records either by full flow
          ($5$-tuple: source IP, source port, destination IP, destination port, protocol) or by
          source/destination IP pair before batching.
    \item \textbf{Batching}: The agent must aggregate records into batches, respecting
          configurable upper bounds on the number of records and the total JSON payload size.
    \item \textbf{Authenticated ingestion}: The backend must accept batched connection records
          only from authenticated clients presenting a valid bearer token.
    \item \textbf{GeoIP enrichment}: The backend must resolve destination IP addresses to
          geographic location (country, city, latitude/longitude) and autonomous system (ASN,
          organisation name) using the MaxMind GeoLite2 databases.
    \item \textbf{Persistence}: The backend must upsert enriched connection records into
          MongoDB, partitioned by tenant.
    \item \textbf{Real-time broadcasting}: The backend must push live connection updates and
          periodic statistics snapshots to the dashboard via Socket.IO.
    \item \textbf{REST API}: The backend must expose REST endpoints for authentication, tenant
          management, connection ingestion, and statistics queries.
    \item \textbf{Multi-tenancy}: Each user may own multiple tenants; data is always scoped to
          a single tenant and never visible across tenant boundaries.
    \item \textbf{Dashboard visualisation}: The frontend must render a world map, a paginated
          connection list, and statistical charts, all updating in real time.
    \item \textbf{Domain DSL}: The backend must load an optional YAML domain DSL at startup to
          configure allowed protocols, ingestion rules, and analytics query parameters without
          code changes.
\end{enumerate}

\section{Non-Functional Requirements}

\begin{itemize}
    \item \textbf{Performance}: The backend should handle continuous ingestion from multiple
          agents without perceptible latency on the dashboard.
    \item \textbf{Security}: All \texttt{/api/*} and ingestion endpoints must be protected by
          bearer JWT authentication; tokens for the agent are short lived (default 12 h).
    \item \textbf{Portability}: The server side components must run on any platform via Docker;
          the capture agent is Linux only due to libpcap dependency.
    \item \textbf{Observability}: The system must expose a \texttt{/health} endpoint and
          publish code coverage reports in CI.
    \item \textbf{Maintainability}: The monorepo must support independent versioning and
          testing of each package via pnpm workspaces and Turborepo.
    \item \textbf{Scalability}: The tenant partitioned data model must accommodate multiple
          monitored hosts without schema changes.
\end{itemize}

% =========================================================
\chapter{Architecture and Design}
\label{ch:design}
% =========================================================

\section{System Overview}

The system is structured as a monorepo with the following top-level packages:

\begin{center}
\begin{tabular}{lll}
\toprule
\textbf{Package} & \textbf{Language} & \textbf{Role} \\
\midrule
\texttt{apps/client-go}   & Go              & Packet capture agent \\
\texttt{apps/backend}     & TypeScript / Node.js & REST API + ingestion + Socket.IO \\
\texttt{apps/dashboard}   & TypeScript / Vue.js~3 & SPA dashboard \\
\texttt{packages/shared}  & TypeScript      & Shared Mongoose models and helpers \\
\texttt{packages/typescript} & N/A          & Shared TypeScript base configuration \\
\bottomrule
\end{tabular}
\end{center}

Figure~\ref{fig:arch} shows the high-level deployment topology. Traefik acts as a single
public entry point and routes requests based on path prefixes: \texttt{/api},
\texttt{/auth}, \texttt{/socket.io}, and \texttt{/health} go to the backend container, while
everything else is served by the dashboard's nginx container.

\begin{figure}[H]
\centering
\resizebox{0.95\textwidth}{!}{%
\begin{tikzpicture}[
    node distance=1.2cm and 1.2cm,
    box/.style={draw, rounded corners=3pt, minimum width=2.6cm, minimum height=0.9cm,
                font=\small, align=center},
    arr/.style={->, thick},
    lbl/.style={font=\scriptsize\itshape}
]

% External
\node[box, fill=gray!15] (client)  {Go Client\\(libpcap)};
\node[box, fill=gray!15, below=0.8cm of client] (browser) {Browser\\(User)};

% Traefik
\node[box, fill=orange!20, right=2.2cm of client, yshift=-0.55cm] (traefik) {Traefik\\(reverse proxy)};

% Backend
\node[box, fill=blue!15, right=2.2cm of traefik, yshift=0.7cm] (backend) {Backend\\Node.js/Express};
\node[box, fill=green!15, right=2.2cm of traefik, yshift=-0.7cm] (dashboard) {Dashboard\\nginx/Vue.js};

% MongoDB
\node[box, fill=yellow!20, right=2.0cm of backend] (mongo) {MongoDB};

% Arrows
\draw[arr] (client)  -- node[above, lbl]{POST /api/connections} (traefik);
\draw[arr] (browser) -- node[below, lbl]{HTTP / WS} (traefik);
\draw[arr] (traefik) -- node[above, lbl]{/api /auth} (backend);
\draw[arr] (traefik) -- node[below, lbl]{/} (dashboard);
\draw[arr] (backend) -- node[above, lbl]{Mongoose} (mongo);
\draw[arr, dashed] (backend) to[bend right=20] node[right, lbl]{Socket.IO} (dashboard);

\end{tikzpicture}%
}
\caption{Deployment topology of the ByteRoute system.}
\label{fig:arch}
\end{figure}

\newpage
\section{Data Flow}

The end-to-end flow of a connection record through the system is as follows.

\begin{enumerate}
    \item The Go client opens a live capture on the specified interface with gopacket/pcap.
          A BPF filter is applied: either user-supplied via \texttt{--bpf} or auto-generated
          from the capture direction and the host's local IPv4 addresses.
    \item Captured packets are decoded into flow records. Depending on the \texttt{--dedupe}
          mode, records are deduplicated by 5-tuple (\texttt{flow}) or by source/destination
          IP pair (\texttt{ip}).
    \item On each flush interval (configurable via \texttt{--flush}), the flow table is
          snapshotted, flows idle longer than \texttt{--idle-ttl} are dropped, and the
          remaining records are split into batches respecting \texttt{--max-batch-conns} and
          \texttt{--max-batch-bytes}.
    \item Each batch is posted to \texttt{POST /api/connections} with a bearer token in the
          \texttt{Authorization} header. The JSON body contains a
          \texttt{reporterIp} field and a \texttt{connections} array.
    \item The backend's Passport.js middleware validates the JWT and resolves the owning
          tenant (from the \texttt{X-Tenant-Id} header, query parameter, or the user's first
          tenant).
    \item The Domain DSL is consulted to filter connections by allowed protocols and apply
          ingestion rules.
    \item For each connection, the backend performs a GeoIP lookup against the local MaxMind
          GeoLite2 databases, resolving city, country, coordinates, ASN, and organisation.
    \item Enriched records are upserted into MongoDB's \texttt{connections} collection,
          partitioned by tenant ID.
    \item The backend emits a \texttt{connections:update} Socket.IO event to all dashboard
          clients subscribed to the affected tenant room.
    \item Periodically (every \texttt{STATS\_EMIT\_INTERVAL} ms, default 30\,s), the backend
          computes and emits aggregated statistics (by country, ASN, and protocol) via
          Socket.IO.
    \item The dashboard receives live events over its persistent Socket.IO connection and
          updates the Pinia stores, which reactively re-render the map, list, and charts.
\end{enumerate}

\section{Backend Internal Architecture}

The backend follows a layered architecture, as depicted in Figure~\ref{fig:backend}.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
    node distance=0.35cm
]
\node[draw, rounded corners=2pt, minimum width=9cm, minimum height=0.75cm,
      font=\small\bfseries, align=center, fill=orange!20] (routes)
      {Routes (\texttt{routes/})};
\node[draw, rounded corners=2pt, minimum width=9cm, minimum height=0.75cm,
      font=\small\bfseries, align=center, fill=blue!10, below=of routes] (ctrl)
      {Controllers (\texttt{controllers/})};
\node[draw, rounded corners=2pt, minimum width=9cm, minimum height=0.75cm,
      font=\small\bfseries, align=center, fill=green!10, below=of ctrl] (svc)
      {Services (\texttt{services/})};
\node[draw, rounded corners=2pt, minimum width=9cm, minimum height=0.75cm,
      font=\small\bfseries, align=center, fill=yellow!10, below=of svc] (infra)
      {Infrastructure: Mongoose models, GeoIP, Auth, DSL (\texttt{infrastructure/})};
\node[draw, rounded corners=2pt, minimum width=9cm, minimum height=0.75cm,
      font=\small\bfseries, align=center, fill=gray!15, below=of infra] (db)
      {MongoDB (via \texttt{packages/shared} Mongoose schemas)};
\end{tikzpicture}
\caption{Layered architecture of the Node.js backend.}
\label{fig:backend}
\end{figure}

\begin{itemize}
    \item \textbf{Routes} (\texttt{src/routes/}): Register Express routers for
          \texttt{/auth}, \texttt{/api/tenants}, \texttt{/api/connections},
          \texttt{/api/metrics}, and \texttt{/health}.
    \item \textbf{Controllers} (\texttt{src/controllers/}): Handle request parsing, input
          validation, and response formation; delegate to services. A dedicated
          \texttt{socket.controller.ts} manages Socket.IO connection lifecycle.
    \item \textbf{Services} (\texttt{src/services/}): Contain business logic for connection
          ingestion, statistics aggregation, and Socket.IO event emission.
    \item \textbf{Infrastructure} (\texttt{src/infrastructure/}): Houses Mongoose model
          definitions (mirroring those in \texttt{packages/shared}), Passport.js JWT
          strategy, the GeoIP lookup wrapper, the Domain DSL compiler/loader, and
          authentication primitives.
    \item \textbf{Middleware} (\texttt{src/middleware/}): Express and Socket.IO middleware
          for authentication, error handling, and tenant resolution.
\end{itemize}

\section{Domain Model}

The persistent entities managed by the system are:

\begin{center}
\begin{tabular}{lp{9cm}}
\toprule
\textbf{Entity} & \textbf{Description} \\
\midrule
\texttt{User}       & An authenticated platform user. Stores hashed credentials and
                       owns one or more tenants. \\
\texttt{Tenant}     & A logical monitoring context (e.g.\ one per host or network
                       segment). All connection and metrics records are scoped to a
                       tenant. \\
\texttt{Connection} & An enriched connection record with source/destination IPs,
                       port, protocol, GeoIP data (country, city, coordinates, ASN,
                       organisation), timestamp, and status. Indexed by tenant ID and
                       destination IP for efficient upsert and query. \\
\bottomrule
\end{tabular}
\end{center}

\section{Security and Multi-Tenancy Design}

Authentication is implemented with Passport.js using a JWT strategy. Two token types exist:

\begin{itemize}
    \item \textbf{User JWT} (default TTL: \texttt{1d}): issued at
          \texttt{POST /auth/signup} and \texttt{POST /auth/signin}. Used by the dashboard and
          for management API calls.
    \item \textbf{Client token} (default TTL: \texttt{12h}): issued at
          \texttt{POST /auth/client-token}. Used exclusively by the Go agent for ingestion, so
          the user's primary credential is never stored on a host.
\end{itemize}

Tenants provide data isolation: every database query includes a \texttt{tenantId} filter, and
the backend refuses to serve or store data for a tenant the authenticated user does not own.
Tenant resolution follows a precedence chain: \texttt{X-Tenant-Id} header, then
\texttt{tenantId} query parameter, then the user's first owned tenant, then the fallback
string \texttt{"default"}.

\section{Domain DSL}

A YAML Domain DSL (\texttt{apps/backend/config/domain.dsl.yaml}) is compiled at startup and
governs two concern areas without requiring code changes:

\begin{itemize}
    \item \texttt{ingestion.connection}: specifies allowed protocols
          (\texttt{TCP}, \texttt{UDP}, \texttt{ICMP}, \texttt{OTHER}), default protocol and
          status, and optionally a list of source IPs to deny.
    \item \texttt{analytics.queries}: controls the sort field, order, and optional record
          limit for each grouped statistics query (\texttt{byCountry}, \texttt{byAsn},
          \texttt{byProtocol}).
\end{itemize}

\section{Go Client Internal Design}

The Go client is organised into internal packages under \texttt{apps/client-go/internal/}:

\begin{center}
\begin{tabular}{ll}
\toprule
\textbf{Package} & \textbf{Responsibility} \\
\midrule
\texttt{capture}  & Opens the pcap handle, applies the BPF filter, and decodes packets
                    into raw flow records. \\
\texttt{flow}     & Maintains the in-memory flow table, handles deduplication and idle
                    eviction. \\
\texttt{backend}  & HTTP client that serialises batches into JSON and posts them to the
                    backend with bearer authentication. \\
\texttt{config}   & Parses CLI flags and environment variables into a unified
                    configuration struct. \\
\texttt{metrics}  & Collects and periodically posts system and capture metrics. \\
\texttt{util}     & Shared utilities (IP parsing, BPF generation, etc.). \\
\bottomrule
\end{tabular}
\end{center}

\section{Dashboard Design}

The Vue.js~3 SPA is structured following the Composition API pattern, with Pinia for state
management and Vue Router for client-side navigation. Key areas of the source tree
(\texttt{apps/dashboard/src/}) include:

\begin{itemize}
    \item \textbf{Views} (\texttt{views/}): Page-level components (map view, connections
          list view, statistics view, login/register).
    \item \textbf{Components} (\texttt{components/}): Reusable UI building blocks (charts,
          tables, tenant selector, header).
    \item \textbf{Stores} (\texttt{stores/}): Pinia stores keyed by feature
          (connections, statistics, authentication); updated reactively by Socket.IO events.
    \item \textbf{Services} (\texttt{services/}): Wrappers around the REST API (Axios
          instance) and the Socket.IO client.
    \item \textbf{Composables} (\texttt{composables/}): Reusable Composition API hooks
          (e.g.\ for tenant resolution, locale formatting).
    \item \textbf{Router} (\texttt{router/}): Route definitions with navigation guards that
          redirect unauthenticated users to the login page.
    \item \textbf{API} (\texttt{api/}): Typed function wrappers for each backend endpoint.
\end{itemize}

The development server (Vite) proxies \texttt{/api}, \texttt{/auth}, and
\texttt{/socket.io} to the backend on port 4000, eliminating CORS issues during development.
In production, Traefik provides the same unified origin.

% =========================================================
\chapter{Technologies}
\label{ch:technologies}
% =========================================================

\section{Go and gopacket/libpcap}

The capture agent is written in Go (version $\geq$~1.26) and relies on the
\texttt{gopacket} library with the \texttt{pcap} backend to perform kernel assisted, BPF filtered
packet capture. Go's concurrency model (goroutines and channels) is used to pipeline packet
decoding and flow-table maintenance without blocking the pcap loop.

\section{Node.js and Express}

The backend targets Node.js with the Express~5 framework. Express provides a minimal,
middleware oriented HTTP server. The \texttt{express.json()} body parser is configured with a
2\,MB limit to match the client's \texttt{--max-batch-bytes} default. A
\texttt{requestTimeout} of 30\,s is set on the underlying \texttt{http.Server} to prevent
stalled connections from blocking the event loop.

\section{Socket.IO}

Socket.IO is layered directly on the same \texttt{http.Server} as Express. It is used
exclusively for server-to-client push: the backend emits \texttt{connections:update} events
when new batches arrive, and aggregated statistics events every 30 seconds. CORS is enabled
for all origins in development (restricted in production by Traefik).

\section{Passport.js and JWT}

Authentication uses Passport.js with the \texttt{passport-jwt} strategy. JWTs are signed with
a configurable HSA256 secret (\texttt{JWT\_SECRET}). Separate token lifetimes apply to user
sessions (\texttt{AUTH\_TOKEN\_TTL}) and client tokens
(\texttt{AUTH\_CLIENT\_TOKEN\_TTL}).

\section{MongoDB and Mongoose}

MongoDB~8 is the persistence layer. Mongoose provides a schema on write ODM. The
\texttt{connections} collection is upserted on each ingestion batch (matching on
\texttt{tenantId} and destination IP) to avoid duplicates. The
\texttt{packages/shared} package centralises Mongoose schemas (\texttt{UserModel},
\texttt{TenantModel}, \texttt{ConnectionModel}) to prevent divergence between the backend and
seeding scripts.

\section{Vue.js 3 and Pinia}

The dashboard is a Vue.js~3 SPA compiled with Vite. The Composition API and
\texttt{<script setup>} syntax are used throughout. Pinia replaces Vuex as the state
management library, providing a simpler, TypeScript friendly API with full devtools support.

\section{Docker Compose and Traefik}

Four services are defined in \texttt{docker-compose.yml}: \texttt{traefik},
\texttt{mongodb}, \texttt{backend}, and \texttt{dashboard}. Traefik~3.6 acts as the single
public entry point on port 8080 (configurable) and routes requests by path prefix using Docker
label rules. MongoDB health checks ensure the backend container only starts once the database
is ready. The backend container downloads GeoLite2 databases automatically on first start via
a Docker entrypoint script.

\section{CI/CD: GitHub Actions and semantic-release}

The CI pipeline (\texttt{.github/workflows/build-and-test.yml}) runs on each push and pull
request: it installs dependencies, builds all packages, runs backend unit and end-to-end
tests, and runs Go tests with coverage. Coverage is merged and published as a Markdown
summary using the \texttt{irongut/CodeCoverageSummary} action. Releases are fully automated
with \texttt{semantic-release} + \texttt{@semantic-release/git}: conventional commits drive
version bumps, changelog updates, and GitHub Release creation.

% =========================================================
\chapter{Implementation Notes}
\label{ch:implementation}
% =========================================================

\section{Client: BPF Filter Generation}

When the user does not supply \texttt{--bpf}, the client introspects the capture interface
to enumerate its IPv4 addresses and constructs a directional BPF expression. For
\texttt{--direction=out} (the default), the filter matches outbound packets:
\texttt{src host <local\_ip>}. This ensures that the agent does not record traffic
originating from other hosts on the same broadcast domain.

\section{Client: Batching and Flush}

The flush loop runs on a configurable ticker. On each tick, the flow table is atomically
snapshotted and cleared. Idle flows (not seen within \texttt{--idle-ttl}) are evicted before
the snapshot. The remaining records are serialised and split at the first boundary that would
exceed \texttt{--max-batch-conns} or \texttt{--max-batch-bytes}, then each batch is posted
sequentially to the backend.

\section{Backend: Ingestion Pipeline}

Code~\ref{lst:ingest} shows a simplified view of the ingestion handler:
the Domain DSL is consulted to filter connections by allowed protocol, GeoIP enrichment is
applied, and records are bulk-upserted via Mongoose.

\begin{lstlisting}[language={}, caption={Simplified ingestion pipeline (TypeScript).}, label=lst:ingest]
async function ingestConnections(ctx, tenantId, body) {
  const { connections, reporterIp } = body;

  // 1. Apply Domain DSL ingestion rules
  const filtered = compiledDsl.filterConnections(connections);

  // 2. GeoIP enrichment
  const enriched = filtered.map((conn) => ({
    ...conn,
    geo: geoipLookup(conn.destIp),
    asn: asnLookup(conn.destIp),
  }));

  // 3. Upsert into MongoDB (partitioned by tenant)
  await ConnectionModel.bulkWrite(
    enriched.map((c) => ({
      updateOne: {
        filter: { tenantId, destIp: c.destIp },
        update: { $set: { ...c, tenantId } },
        upsert: true,
      },
    }))
  );

  // 4. Push live update to connected dashboard clients
  io.to(tenantId).emit("connections:update", enriched);
}
\end{lstlisting}

\section{Backend: Statistics Emission}

Every \texttt{STATS\_EMIT\_INTERVAL} milliseconds the backend runs MongoDB aggregation
pipelines to compute grouped statistics (by country, ASN, and protocol) for every active
tenant, and emits the results as Socket.IO events. The aggregation sort order and optional
limit are driven by the Domain DSL \texttt{analytics.queries} configuration, allowing the
operator to tune result sets without redeployment.

\section{Backend: Environment Configuration}

Table~\ref{tab:env} lists the most important backend environment variables.

\begin{table}[H]
\centering
\begin{tabular}{llp{5.5cm}}
\toprule
\textbf{Variable} & \textbf{Default} & \textbf{Description} \\
\midrule
\texttt{MONGODB\_URI}                  & \texttt{mongodb://mongodb:27017/byteroute} & MongoDB connection string \\
\texttt{JWT\_SECRET}                   & (required)   & HMAC-SHA256 signing key \\
\texttt{PORT}                          & \texttt{4000} & HTTP listening port \\
\texttt{MAXMIND\_USER\_ID}             & N/A          & MaxMind account ID \\
\texttt{MAXMIND\_API\_KEY}             & N/A          & MaxMind license key \\
\texttt{CONNECTIONS\_BOOTSTRAP\_LIMIT} & \texttt{500} & Max connections loaded on startup \\
\texttt{STATS\_EMIT\_INTERVAL}         & \texttt{30000} & Statistics emission period (ms) \\
\texttt{AUTH\_TOKEN\_TTL}              & \texttt{1d}  & User JWT expiry \\
\texttt{AUTH\_CLIENT\_TOKEN\_TTL}      & \texttt{12h} & Client token expiry \\
\bottomrule
\end{tabular}
\caption{Key backend environment variables.}
\label{tab:env}
\end{table}

\section{Client: Environment Variables}

The Go client supports both CLI flags and environment variable equivalents:

\begin{center}
\begin{tabular}{ll}
\toprule
\textbf{Variable} & \textbf{CLI flag} \\
\midrule
\texttt{BYTEROUTE\_BACKEND\_URL}  & \texttt{--backend} \\
\texttt{BYTEROUTE\_AUTH\_TOKEN}   & \texttt{--auth-token} \\
\texttt{BYTEROUTE\_IFACE}         & \texttt{--iface} \\
\texttt{BYTEROUTE\_BPF}           & \texttt{--bpf} \\
\texttt{BYTEROUTE\_REPORTER\_IP}  & \texttt{--reporter-ip} \\
\texttt{BYTEROUTE\_HOST\_ID}      & (host identifier) \\
\bottomrule
\end{tabular}
\end{center}

\section{REST API Reference}

Code~\ref{lst:curl} demonstrates the standard authentication and setup workflow using
\texttt{curl}, showing the exact sequence a new user would follow to start ingesting data.

\begin{lstlisting}[language=bash, caption={Authentication and setup via cURL.}, label=lst:curl]
# 1. Register and obtain a user JWT
AUTH_TOKEN=$(curl -sS -X POST http://localhost:8080/auth/signup \
  -H 'content-type: application/json' \
  -d '{"name":"Demo","email":"demo@byteroute.dev","password":"secret"}' \
  | jq -r '.token')

# 2. Create a tenant
curl -sS -X POST http://localhost:8080/api/tenants \
  -H "authorization: Bearer $AUTH_TOKEN" \
  -H 'content-type: application/json' \
  -d '{"name":"My Laptop"}'

# 3. Generate a short-lived client token for the Go agent
CLIENT_TOKEN=$(curl -sS -X POST http://localhost:8080/auth/client-token \
  -H "authorization: Bearer $AUTH_TOKEN" \
  | jq -r '.token')

# 4. Start the capture agent
sudo ./byteroute-client \
  --iface eth0 \
  --backend http://localhost:8080 \
  --auth-token "$CLIENT_TOKEN" \
  --flush 5s
\end{lstlisting}

Table~\ref{tab:api} summarises the full API surface.

\begin{table}[H]
\centering
\begin{tabular}{llp{5.5cm}}
\toprule
\textbf{Method \& Path} & \textbf{Auth} & \textbf{Description} \\
\midrule
\texttt{POST /auth/signup}        & No  & Create account; returns user JWT \\
\texttt{POST /auth/signin}        & No  & Authenticate; returns user JWT \\
\texttt{GET  /auth/me}            & Yes & Current user info \\
\texttt{POST /auth/client-token}  & Yes & Issue a short-lived agent token \\
\texttt{GET  /api/tenants}        & Yes & List tenants owned by user \\
\texttt{POST /api/tenants}        & Yes & Create a new tenant \\
\texttt{POST /api/connections}    & Yes & Ingest a batch of connections \\
\texttt{POST /api/metrics}        & Yes & Ingest metrics snapshots \\
\texttt{GET  /health}             & No  & Returns \texttt{200 OK} when ready \\
\bottomrule
\end{tabular}
\caption{Backend REST API endpoints.}
\label{tab:api}
\end{table}

\section{Ingestion Payload Schema}

The client posts the following JSON structure to \texttt{POST /api/connections}:

\begin{lstlisting}[language={}, caption={Connection ingestion payload.}, label=lst:payload]
{
  "reporterIp": "203.0.113.10",
  "connections": [
    {
      "id": "tcp-192.0.2.5:443-203.0.113.10:52301",
      "sourceIp": "192.0.2.5",
      "sourcePort": 443,
      "destIp": "203.0.113.10",
      "destPort": 52301,
      "protocol": "TCP"
    }
  ]
}
\end{lstlisting}

The backend responds with HTTP~202 Accepted after enqueuing the enrichment and upsert
operations.

% =========================================================
\chapter{Testing and Validation}
\label{ch:testing}
% =========================================================

\section{Backend Testing Strategy}

The backend test suite is located in \texttt{apps/backend/test/} and is executed with Vitest.
Two test categories are distinguished:

\begin{itemize}
    \item \textbf{Unit tests}: cover individual services, middleware, utility functions, and
          domain logic in isolation. External dependencies (MongoDB, MaxMind) are replaced
          with mocks defined in \texttt{test/mocks/}.
    \item \textbf{End-to-end (E2E) tests}: spin up the full Express application (including
          a real MongoDB instance and an in process Socket.IO server) and exercise complete
          request-response cycles. Test fixtures cover authentication flows, connection
          ingestion, tenant creation, and statistics emission. The E2E suite is located in
          \texttt{test/features/}.
\end{itemize}

Scripts in \texttt{apps/backend/scripts/} handle coverage merging between the two test
runs (\texttt{merge-coverage.mjs}) so that a unified coverage report is produced for the
GitHub Actions job summary.

\section{Go Client Testing}

Go unit tests accompany each internal package. The coverage output is produced with
\texttt{go test -coverprofile=coverage.out ./...} and converted to Cobertura XML for
consumption by the \texttt{irongut/CodeCoverageSummary} action.

\section{CI Coverage Reporting}

The GitHub Actions workflow merges backend (Vitest V8) and Go coverage reports and publishes
a line-coverage summary as a job annotation. Artifacts uploaded per run include:

\begin{itemize}
    \item \texttt{coverage-reports}: HTML coverage report and
          \texttt{code-coverage-results.md} for the backend.
    \item \texttt{coverage-go}: raw \texttt{coverage.out} and Cobertura XML for Go.
\end{itemize}

\section{Integration Validation}

The system was validated end-to-end by:

\begin{enumerate}
    \item Deploying the full stack with \texttt{docker compose up -d}.
    \item Verifying \texttt{GET /health} returns \texttt{200 OK}.
    \item Registering a user and creating a tenant via the dashboard UI.
    \item Generating a client token and launching the Go agent against a local interface.
    \item Observing real-time connection updates appearing on the live map and connection
          list in the dashboard.
    \item Confirming GeoIP enrichment by checking that known public IP destinations were
          resolved to their correct country and ASN.
\end{enumerate}

% =========================================================
\chapter{Deployment}
\label{ch:deployment}
% =========================================================

\section{Prerequisites}

\begin{itemize}
    \item Docker and Docker Compose.
    \item A free MaxMind account to obtain a \texttt{MAXMIND\_USER\_ID} and
          \texttt{MAXMIND\_API\_KEY} (required for GeoLite2 database download at container
          startup).
    \item On the capture host: Linux with  libpcap headers, or the ability to run the
          pre-built Docker image with \texttt{--network host} and the
          \texttt{cap\_net\_raw,cap\_net\_admin} Linux capabilities.
\end{itemize}

\section{Docker Compose Quickstart}

\begin{lstlisting}[language=bash, caption={Production deployment with Docker Compose.}, label=lst:deploy]
# 1. Create .env in the repo root
cat > .env <<EOF
MAXMIND_USER_ID=<your_id>
MAXMIND_API_KEY=<your_key>
JWT_SECRET=$(openssl rand -hex 32)
EOF

# 2. Start the stack (Traefik + MongoDB + backend + dashboard)
docker compose up -d

# 3. Verify
curl http://localhost:8080/health   # should return 200 OK
\end{lstlisting}

The backend Docker entrypoint (\texttt{docker-entrypoint.sh}) runs a MaxMind download
script before starting the Node.js process, ensuring the GeoLite2 databases are present at
runtime.

\section{Development Setup}

From the repository root (requires Node.js $\geq$~the version pinned in
\texttt{package.json} and pnpm):

\begin{lstlisting}[language=bash, caption={Local development setup.}, label=lst:devsetup]
pnpm install           # install all workspaces
pnpm -F @byteroute/backend dev      # backend on :4000
pnpm -F @byteroute/dashboard dev    # dashboard on :3000
\end{lstlisting}

For the Go client (requires Go $\geq$~1.26 and \texttt{libpcap-dev}):

\begin{lstlisting}[language=bash, caption={Building and running the Go client from source.}, label=lst:gobuild]
cd apps/client-go
go mod tidy
go build -o byteroute-client ./cmd/byteroute-client

# Run with elevated privileges
sudo ./byteroute-client \
  --iface eth0 \
  --backend http://localhost:4000 \
  --auth-token "$CLIENT_TOKEN" \
  --flush 5s
\end{lstlisting}

\section{CI/CD and Release Process}

On every push to the default branch that contains conventional commits, \texttt{semantic-release}:

\begin{enumerate}
    \item Analyses commit messages to determine the next semantic version.
    \item Updates \texttt{CHANGELOG.md} and workspace \texttt{package.json} files.
    \item Creates a tagged GitHub Release with auto-generated release notes.
    \item Triggers Docker image builds for \texttt{byteroute-backend} and
          \texttt{byteroute-dashboard}, which are pushed to the GitHub Container
          Registry (\texttt{ghcr.io/ste-script/}).
\end{enumerate}

% =========================================================
\chapter{Conclusion}
\label{ch:conclusion}
% =========================================================

ByteRoute delivers a complete, end-to-end solution for real-time network traffic monitoring
and geo-visualisation. The architecture cleanly separates concerns across three independently
deployable components (Go capture agent, Node.js/Express backend, and Vue.js~3 dashboard),
connected through well defined REST and WebSocket contracts.

Key design strengths include the multi-tenancy model, which allows a single backend instance
to serve multiple monitored hosts; the Domain DSL, which externalises configuration concerns
such as protocol filtering and statistics shaping; and the short-lived client token, which
decouples the capture agent's credential lifetime from the user session.

The testing strategy combines Vitest unit and end-to-end tests for the backend with Go
coverage, all integrated in a single GitHub Actions workflow that publishes reproducible
coverage summaries and enforces a quality gate on every pull request.

Possible directions for future work include:

\begin{itemize}
    \item \textbf{Cross-platform client}: porting the capture agent to macOS and Windows
          using platform-appropriate packet-capture APIs.
    \item \textbf{Enhanced enrichment}: integrating additional sources such as RDAP or
          threat-intelligence feeds to label connections as potentially malicious.
    \item \textbf{Alert rules}: allowing users to configure threshold based or
          pattern based alerts with dashboard notifications or email delivery.
    \item \textbf{Horizontal scalability}: replacing the single in process Socket.IO
          server with a Redis adapter to support multiple backend instances behind
          a load balancer.
\end{itemize}

\end{document}
