\documentclass[12pt]{report}
\usepackage{xcolor}

\title{\textbf{ByteRoute} \\ \large SPE Report}

\author{
  Babini Stefano\\
  0001125127 \\
  \texttt{stefano.babini5@studio.unibo.it}
}

\date{\today}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage[english]{babel}
\usepackage{float}
\pagestyle{plain}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage[export]{adjustbox}
\usetikzlibrary{petri,positioning,shapes.geometric}
\usepackage{listings}
\usepackage{titlesec}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{placeins}

\titleformat{\chapter}[display]{\normalfont\bfseries}{}{0pt}{\Large}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    filecolor=magenta,
    urlcolor=cyan,
}

\graphicspath{{images/}}

\renewcommand{\lstlistingname}{Code}
\renewcommand{\lstlistlistingname}{List of \lstlistingname s}

% Color setup
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

% Code listing style
\lstdefinestyle{mystyle}{
  backgroundcolor=\color{backcolour},
  commentstyle=\color{codegreen},
  keywordstyle=\color{magenta},
  numberstyle=\tiny\color{codegray},
  stringstyle=\color{codepurple},
  basicstyle=\ttfamily\footnotesize,
  breakatwhitespace=false,
  breaklines=true,
  captionpos=b,
  keepspaces=true,
  numbers=left,
  numbersep=5pt,
  showspaces=false,
  showstringspaces=false,
  showtabs=false,
  tabsize=2
}

\lstset{style=mystyle}

\begin{document}

\maketitle
\newpage
\tableofcontents{}
\newpage

% ============================================================
\chapter{Introduction}
% ============================================================

This report documents the design, architecture, implementation, and development
process of \textbf{ByteRoute}, a network traffic monitoring platform developed
as a university project for the SPE (Software Performance Engineering) and ASW
(Architetture Software per il Web) curricula at the University of Bologna.

ByteRoute enables users to gain real-time visibility into the internet traffic
produced by their Linux hosts. By capturing network packets at the operating
system level, enriching connection records with geographic and organizational
metadata, and presenting the results through an interactive web dashboard, the
system answers a concrete operational need: understanding which remote servers
and autonomous systems a machine communicates with at any given moment.

\section{Motivation}

Modern network activity is complex and often opaque. Desktop and server
machines establish dozens of concurrent connections to third-party services,
content delivery networks, and cloud providers. Without dedicated tooling,
identifying those destinations requires either deep packet inspection expertise
or expensive commercial software. ByteRoute addresses this gap with a
lightweight, open-source, and self-hosted solution that any developer or system
administrator can deploy with Docker Compose.

\newpage
\section{Report Structure}

The remainder of this document is organized as follows:

\begin{itemize}
  \item Chapter~\ref{chap:problem} presents the problem analysis, the proposed
    solution, and the involved user personas and actors.
  \item Chapter~\ref{chap:requirements} enumerates functional and non-functional
    requirements.
  \item Chapter~\ref{chap:domain} defines the domain model and the ubiquitous
    language.
  \item Chapter~\ref{chap:architecture} describes the physical and logical
    architecture.
  \item Chapter~\ref{chap:implementation} details the implementation of each
    component.
  \item Chapter~\ref{chap:devops} covers the DevOps pipeline.
  \item Chapter~\ref{chap:validation} reports on testing and validation results.
  \item Chapter~\ref{chap:conclusion} concludes with a summary and future
    directions.
\end{itemize}

% ============================================================
\chapter{Problem Analysis}
\label{chap:problem}
% ============================================================

\section{Context}

Network monitoring is a well established field, yet most available tools target
either professional network engineers (with complex configuration overhead) or
enterprise environments (with licensing costs). Individual developers and small
teams running Linux workstations or home servers have limited access to
user-friendly, self-hosted monitoring that shows, in real time, the geographic
and organizational distribution of their outbound traffic.

\section{Problem Statement}

The absence of accessible, real-time network traffic dashboards leads to several
practical difficulties:

\begin{enumerate}
  \item \textbf{Lack of visibility}: there is no simple way to observe which
    Autonomous Systems (ASNs) or countries a host is contacting at a given
    moment.
  \item \textbf{Fragmented tooling}: existing options such as \texttt{tcpdump}
    or \texttt{Wireshark} require specialized knowledge and do not provide
    persistent, aggregated, or geographically enriched records.
  \item \textbf{No multi-host support}: administrators managing several machines
    cannot correlate their traffic in a single view without custom
    infrastructure.
\end{enumerate}

\newpage
\section{Proposed Solution}

ByteRoute solves these problems through a three-tier architecture:

\begin{enumerate}
  \item A \textbf{lightweight Linux client} written in Go captures packets from
    a network interface, aggregates them into connection-level records,
    deduplicates them, and forwards batches to the backend over HTTPS.
  \item A \textbf{Node.js backend} receives the connection batches, enriches
    each record with GeoIP metadata (country, city, latitude/longitude, ASN,
    organization name) from MaxMind GeoLite2 databases, persists the records in
    MongoDB, and broadcasts live updates to connected dashboard clients via
    Socket.IO.
  \item A \textbf{Vue.js 3 dashboard} presents a world map of traffic flows, a
    live connection list, and statistics panels grouped by country, ASN, and
    protocol.
\end{enumerate}

The entire stack is containerized with Docker and orchestrated with Docker
Compose, keeping deployment reproducible and independent of the host
operating system.

% ============================================================
\chapter{Requirements Analysis}
\label{chap:requirements}
% ============================================================

\section{System Requirements}

\begin{enumerate}
  \item \textbf{Packet capture}: the client must intercept IP-level network
    traffic on a specified Linux interface using libpcap, without modifying
    kernel settings beyond the required capabilities (\texttt{cap\_net\_raw},
    \texttt{cap\_net\_admin}).
  \item \textbf{Connection enrichment}: the backend must annotate each
    connection record with geographic coordinates, country code, ASN, and
    organization name sourced from the MaxMind GeoLite2 databases.
  \item \textbf{Persistent storage}: enriched connection records and traffic
    metrics must be stored in a MongoDB collection, organized per tenant, and
    queryable for historical statistics.
  \item \textbf{Real-time propagation}: state changes (new connections, updated
    statistics) must be propagated to connected dashboard clients within seconds,
    using a push model based on WebSockets.
  \item \textbf{Multi-tenancy}: connection data must be isolated per tenant so
    that different hosts or organizational units do not share records.
  \item \textbf{Containerization}: all services must be deployable via Docker
    Compose and must not depend on host-specific configuration beyond the
    provided environment files.
\end{enumerate}

\newpage
\section{Functional Requirements}

\begin{enumerate}
  \item As a user, I want to see a world map that shows where my machine's
    outbound connections are directed, with pins or flow lines indicating
    destination countries.
  \item As a user, I want to view a live list of active connections, each
    showing the destination IP, country, ASN, organization name, and protocol.
  \item As a user, I want statistics panels that aggregate connections by
    country, by ASN, and by protocol, updated in real time.
  \item As a user, I want to switch between tenants to observe the traffic of
    different monitored hosts.
  \item As an account owner, I want to register an account, sign in, and receive a JWT
    token for subsequent authenticated requests.
  \item As an account owner, I want to create tenants and generate client tokens for
    the Go client, so that the capture agent can authenticate its ingestion
    requests.
  \item As a Go client operator, I want to specify the capture interface,
    direction filter, BPF expression, deduplication strategy, batch size,
    and flush interval on the command line or via environment variables.
\end{enumerate}

\newpage
\section{Non-Functional Requirements}

\begin{enumerate}
  \item \textbf{Performance}: the Go client must handle sustained packet rates
    without dropping events; aggregation and deduplication must occur in memory
    before each batch is sent to the backend.
  \item \textbf{Security}: all \texttt{/api/*} routes must be protected by
    bearer authentication; client tokens must carry an expiration time and must
    be separate from user session tokens.
  \item \textbf{Portability}: the Go client targets Linux only (libpcap
    dependency); all other components must run on any platform supported by the
    container runtime.
  \item \textbf{Maintainability}: the project must be organized as a monorepo
    with clear per-component boundaries, shared typed contracts, and automated
    dependency updates via Renovate.
  \item \textbf{Observability}: the coverage pipeline must publish machine-readable
    Cobertura XML and human-readable HTML reports as CI artifacts after every
    test run.
  \item \textbf{Configurability}: ingestion rules and analytics query parameters
    must be externalized in a YAML domain DSL, allowing behaviour changes without
    modifying source code.
\end{enumerate}

% ============================================================
\chapter{Domain Model}
\label{chap:domain}
% ============================================================

This chapter defines the conceptual model of the ByteRoute domain, establishing
the ubiquitous language shared between the client, backend, and dashboard
components.

\section{Ubiquitous Language}

Table~\ref{tab:ubiquitous} lists the core terms used throughout the codebase
and this report.

\begin{table}[htbp]
  \centering
  \small
  \setlength{\tabcolsep}{4pt}
  \renewcommand{\arraystretch}{0.95}
  \caption{Ubiquitous language for the ByteRoute domain.}
  \label{tab:ubiquitous}
  \begin{tabularx}{\textwidth}{p{0.22\textwidth}X}
    \toprule
    \textbf{Term} & \textbf{Definition} \\
    \midrule
    \textbf{Connection} &
      A record representing a single observed network flow between a source IP
      and a destination IP, annotated with protocol, port, timestamps, byte
      counts, and GeoIP metadata. This is the primary aggregate of the
      domain. \\
    \addlinespace
    \textbf{PacketEvent} &
      A raw event emitted by the Go capture layer containing the source IP,
      destination IP, ports, protocol, and packet length extracted from a
      single captured IP datagram. \\
    \addlinespace
    \textbf{Flow} &
      The in-memory, deduplicated accumulation of packet events that share the
      same five-tuple (or IP pair, depending on deduplication strategy). Flows
      are produced by the Go client and converted to Connection records before
      being sent to the backend. \\
    \addlinespace
    \textbf{Tenant} &
      A logical namespace that isolates connection data for a specific user or
      monitored host. Each connection record is tagged with a tenant identifier,
      and all queries are scoped accordingly. \\
    \addlinespace
    \textbf{GeoIpEnrichment} &
      A value object attached to each Connection after backend processing,
      containing the destination country, country code, city, latitude,
      longitude, ASN number, and AS organization name. \\
    \addlinespace
    \textbf{TrafficFlow} &
      A lightweight projection of a Connection used by the dashboard to render
      map arcs. It carries source and destination coordinates and the connection
      count. \\
    \addlinespace
    \textbf{Snapshot} &
      A metric record capturing aggregated statistics at a point in time, used
      by the metrics ingestion endpoint and the statistics panels. \\
    \addlinespace
    \textbf{Token} &
      A signed JWT containing the user identity (identifier and email, with an
      optional display name), the authorized tenant identifiers, and an
      expiration time. Two token types exist: the user session token (TTL of 1
      day by default) and the client token (TTL of 12 hours by default). \\
    \addlinespace
    \textbf{Domain DSL} &
      A YAML configuration file (default: \texttt{config/domain.dsl.yaml})
      compiled at backend startup. It governs ingestion rules (allowed
      protocols, denied source IPs) and analytics query parameters (sort order,
      limit) without requiring code changes. \\
    \bottomrule
  \end{tabularx}
\end{table}

\FloatBarrier

\section{Domain Structure}

The backend domain is organized into three sub-domains, each residing under
\texttt{apps/backend/src/domain/}:

\begin{itemize}
  \item \texttt{connection/}: defines the \texttt{Connection} type, the
    \texttt{GeoIpEnrichment} value object, the \texttt{IngestResult} type, and
    the interfaces for the connection repository and the GeoIP service. This
    sub-domain is the core of the system.
  \item \texttt{metrics/}: defines the snapshot types used by the metrics
    ingestion pipeline.
  \item \texttt{identity/}: defines the principal type representing the
    authenticated caller, used by the middleware layer to propagate identity
    across request handlers.
\end{itemize}

The \texttt{packages/shared} package acts as a shared kernel: it exports the
\texttt{Connection} and \texttt{TrafficFlow} types together with the Mongoose
models, guaranteeing that the TypeScript type definitions used in the backend
services are identical to those referenced in the dashboard's API layer.

% ============================================================
\chapter{Architecture and Design}
\label{chap:architecture}
% ============================================================

\section{Physical Architecture}

ByteRoute is a client-server system executed inside Docker containers and
coordinated through Docker Compose. Figure~\ref{fig:physical} illustrates the
high-level deployment topology.

\begin{figure}[H]
  \centering
  \begin{adjustbox}{max width=\textwidth, max totalheight=0.75\textheight, keepaspectratio}
  \begin{tikzpicture}[
    box/.style={rectangle, draw, rounded corners=4pt,
                minimum width=2.8cm, minimum height=1cm,
                align=center, font=\small},
    dashed box/.style={rectangle, draw, dashed, rounded corners=4pt,
                       minimum width=8cm, minimum height=3.8cm,
                       align=center, font=\small\itshape},
    arr/.style={->, >=stealth, thick},
    node distance=1.6cm and 2.2cm
  ]
    % Linux host (outside containers)
    \node[box, fill=backcolour] (client)
      {\textbf{Go Client}\\libpcap};

    % Docker Compose group
    \node[box, fill=backcolour, right=2.6cm of client] (traefik)
      {\textbf{Traefik}\\Reverse Proxy};

    \node[box, fill=backcolour, right=2.0cm of traefik] (backend)
      {\textbf{Backend}\\Node.js / Express};

    \node[box, fill=backcolour, below=1.4cm of traefik] (dashboard)
      {\textbf{Dashboard}\\Vue.js 3 / Nginx};

    \node[box, fill=backcolour, right=2.0cm of backend] (mongo)
      {\textbf{MongoDB}};

    % Arrows
    \draw[arr] (client) -- node[above, font=\scriptsize]{HTTPS POST} (traefik);
    \draw[arr] (traefik) -- node[above, font=\scriptsize]{proxy} (backend);
    \draw[arr] (traefik) -- node[left, font=\scriptsize]{serve} (dashboard);
    \draw[arr] (backend) -- node[above, font=\scriptsize]{read / write} (mongo);
    \draw[arr, <->] (dashboard.east)
      -- ++(0.6,0)
      |- node[right, font=\scriptsize, pos=0.3, align=left]{Socket.IO\\+ HTTP} (backend.south);
  \end{tikzpicture}
  \end{adjustbox}
  \caption{Physical deployment topology of ByteRoute.}
  \label{fig:physical}
\end{figure}

The four deployed services are:

\begin{itemize}
  \item \textbf{Go Client}: runs on the monitored Linux host (or inside a
    privileged container). It accesses the raw network interface via libpcap
    and pushes connection batches to the backend over HTTPS.
  \item \textbf{Backend (Node.js/Express)}: the central service. It handles
    authentication, connection ingestion, GeoIP enrichment, MongoDB persistence,
    and real-time broadcasts to dashboard clients via Socket.IO.
  \item \textbf{Dashboard (Vue.js 3 / Nginx)}: a Single Page Application
    served by Nginx inside a Docker container. It communicates with the backend
    through HTTP REST calls and a persistent Socket.IO connection.
  \item \textbf{MongoDB}: a NoSQL database instance that stores connection
    records, user accounts, and tenant definitions.
\end{itemize}

Traefik acts as the ingress reverse proxy, routing browser and client
requests to the appropriate backend or dashboard container by host and path
rules. In development, the Vite dev server provides an equivalent proxy for
\texttt{/api}, \texttt{/auth}, and \texttt{/socket.io} paths.

\section{Logical Architecture}

Within the backend, the codebase follows a layered organization inspired by
Hexagonal Architecture (Ports and Adapters). The three layers are:

\begin{enumerate}
  \item \textbf{Domain layer}: contains type definitions, interfaces (ports),
    and pure business logic. It has no dependency on Express, MongoDB, or any
    external library. The \texttt{normalizeConnection} function and the
    \texttt{applyIngestionDsl} function both reside in this layer or at the
    boundary to the application layer.
  \item \textbf{Application/Service layer}: contains the services that
    orchestrate domain logic. Notable services include \texttt{ingest.ts}
    (connection ingestion pipeline), \texttt{geoip.ts} (enrichment
    coordination), \texttt{auth.service.ts} (JWT issuance and verification),
    and \texttt{tenant.service.ts} (tenant resolution and management).
  \item \textbf{Infrastructure layer}: contains Express route handlers,
    Mongoose models, the Socket.IO server wrapper, and the compiled domain DSL
    adapter. This layer depends on all external frameworks and libraries.
\end{enumerate}

The repository is organized as a \texttt{pnpm} monorepo using Turbo for task
orchestration. The top-level layout is:

\begin{itemize}
  \item \texttt{apps/backend}: the Node.js backend application;
  \item \texttt{apps/dashboard}: the Vue.js 3 frontend application;
  \item \texttt{apps/client-go}: the Go capture client;
  \item \texttt{packages/shared}: shared TypeScript types and Mongoose models;
  \item \texttt{packages/typescript}: shared TypeScript base configuration;
  \item \texttt{scripts}: workspace-level automation and documentation scripts.
\end{itemize}

\section{Data Flow}

The complete ingestion-to-visualization pipeline proceeds through the following
stages:

\begin{enumerate}
  \item The Go client opens a libpcap handle on a network interface and emits
    \texttt{PacketEvent} values for each captured IP datagram.
  \item The \texttt{Aggregator} (inside \texttt{internal/flow}) groups events
    by five-tuple key (or by IP pair, depending on the \texttt{--dedupe} flag),
    updates per-flow byte and packet counters, and marks flows as dirty when
    they receive new traffic.
  \item At the end of each flush interval, the aggregator exports a batch of
    dirty flows, serializes them to a JSON payload, and sends an HTTP POST
    request to \texttt{/api/connections} on the backend.
  \item The backend's ingestion service applies the domain DSL ingestion rules
    (filtering denied source IPs, normalizing protocols), then calls
    \texttt{enrichBatch} to annotate each record with GeoIP metadata.
  \item Enriched records are upserted into MongoDB using bulk update operations
    (filter on \texttt{tenantId} and \texttt{id}).
  \item The service calls \texttt{emitTrafficFlows} and
    \texttt{emitStatisticsUpdate} on the Socket.IO server, which broadcasts the
    new data to all dashboard clients subscribed to the relevant tenant room.
  \item The Vue.js dashboard receives the Socket.IO events and reactively
    updates the world map pins, the connection list, and the statistics panels
    without a page reload.
\end{enumerate}

% ============================================================
\chapter{Implementation Notes}
\label{chap:implementation}
% ============================================================

\section{Technologies}

Table~\ref{tab:tech} lists the primary technologies used in each component.

\begin{table}[H]
  \centering
  \caption{Technology stack of ByteRoute.}
  \label{tab:tech}
  \begin{tabularx}{\textwidth}{llX}
    \toprule
    \textbf{Component} & \textbf{Language} & \textbf{Key dependencies} \\
    \midrule
    Go Client &
      Go ($\geq$ 1.26) &
      \texttt{gopacket/pcap} (libpcap bindings), standard library
      (\texttt{net/http}, \texttt{sync}, \texttt{encoding/json}) \\
    \addlinespace
    Backend &
      TypeScript / Node.js &
      Express, Socket.IO, Mongoose, \texttt{jsonwebtoken}, \texttt{bcryptjs},
      \texttt{maxmind} (GeoLite2 reader), Vitest \\
    \addlinespace
    Dashboard &
      TypeScript / Vue.js 3 &
      Vite, Pinia, Vue Router, \texttt{socket.io-client}, Vitest, Nginx
      (production serving) \\
    \addlinespace
    Shared Package &
      TypeScript &
      Mongoose schema definitions, domain type exports \\
    \addlinespace
    Infrastructure &
      YAML / Shell &
      Docker, Docker Compose, Traefik, GitHub Actions, Renovate \\
    \bottomrule
  \end{tabularx}
\end{table}

\section{Backend}

\subsection{Connection Ingestion Pipeline}

The ingestion entry point is exposed by the Express router at
\texttt{POST /api/connections}. After authentication middleware validates the
bearer token and extracts the tenant identifier, the route handler delegates
to \texttt{ingestConnections} in \texttt{src/services/ingest.ts}.

The function first applies ingestion rules compiled from the domain DSL.
Code~\ref{lst:dsl-apply} shows the relevant logic.

\begin{lstlisting}[language=TypeScript, caption={Domain DSL application during ingestion.}, label={lst:dsl-apply}]
function applyIngestionDsl(connections: Connection[]): Connection[] {
  const rules = getCompiledDomainDsl().ingestion.connection;

  return connections
    .filter((c) => !rules.denySourceIps.has(c.sourceIp))
    .map((c) => {
      const protocol = rules.allowedProtocols.has(c.protocol)
        ? c.protocol
        : rules.defaultProtocol;
      const status = c.status ?? rules.defaultStatus;
      return { ...c, protocol, status };
    });
}
\end{lstlisting}

After filtering, the service calls \texttt{enrichBatch}, which reads
MaxMind GeoLite2 databases (ASN and City) from disk and appends geographic
and organizational metadata to each connection record. The enriched batch is
then upserted into MongoDB using a bulk \texttt{updateOne} with
\texttt{upsert: true} for each record, keyed on \texttt{(tenantId, id)}.
Finally, the service emits updated traffic-flow projections and statistics
through Socket.IO to dashboard clients watching the relevant tenant room.

\subsection{Authentication and Token Management}

The backend exposes four authentication routes under \texttt{/auth}:
\texttt{POST /auth/signup}, \texttt{POST /auth/signin}, \texttt{GET /auth/me},
and \texttt{POST /auth/client-token}. Passwords are hashed with a
\texttt{bcryptjs}-based utility (the \texttt{password.ts} service) before
storage, and never persisted in plain text.

On successful sign-in, the backend issues a signed JWT containing the user
identifier, email, and the set of tenant identifiers the user can access. A
second, short-lived token type (the
\textit{client token}) is issued separately via \texttt{/auth/client-token} and
is intended for use by the Go client. Both token types carry an expiration time
configurable via environment variables (\texttt{AUTH\_TOKEN\_TTL} and
\texttt{AUTH\_CLIENT\_TOKEN\_TTL}).

All \texttt{/api/*} routes are protected by a Passport.js bearer strategy that
verifies the token signature and expiration before injecting the principal into
the request context.

\subsection{Multi-Tenancy}

Tenant resolution is explicit: the tenant identifier must be provided by the
client on each request (via the \texttt{X-Tenant-Id} HTTP header or the
\texttt{tenantId} query parameter). If no tenant identifier is supplied, the
backend rejects the request as \texttt{401 Unauthorized}. The authenticated
principal is then checked for access to the requested tenant.

For real-time connections, the Socket.IO handshake must likewise include a
tenant identifier (via \texttt{handshake.auth.tenantId}, \texttt{handshake.query.tenantId},
or the \texttt{X-Tenant-Id} header); missing tenant identifiers are rejected
as unauthorized.

\subsection{Real-Time Events via Socket.IO}

The backend maintains a Socket.IO server that operates over the same HTTP
server as Express. When a client connects, it authenticates with the same
bearer token and joins a Socket.IO room named after its tenant identifier.
This room-based isolation ensures that traffic events and statistics updates
emitted for tenant A are never delivered to dashboard sessions authorized for
tenant B.

Two distinct emission events are used: \texttt{trafficFlows} (a minimal
projection carrying source and destination coordinates for map rendering) and
\texttt{statisticsUpdate} (aggregated counts grouped by country, ASN, and
protocol, governed by the domain DSL analytics rules).

\subsection{Domain DSL}

The domain DSL is a YAML file compiled once at startup by
\texttt{src/infrastructure/dsl/domain-dsl.ts}. Code~\ref{lst:dsl-yaml} shows
the default configuration.

\begin{lstlisting}[language=bash, caption={Default domain DSL configuration (\texttt{config/domain.dsl.yaml}).}, label={lst:dsl-yaml}]
ingestion:
  connection:
    allowedProtocols: [TCP, UDP, ICMP, OTHER]
    defaultProtocol: OTHER
    defaultStatus: active
    denySourceIps: []

analytics:
  queries:
    byCountry:
      sortBy: connections
      order: desc
    byAsn:
      sortBy: connections
      order: desc
    byProtocol:
      sortBy: connections
      order: desc
\end{lstlisting}

The \texttt{ingestion.connection} section controls which protocols are accepted
and which source IPs are denied before enrichment. The
\texttt{analytics.queries} section shapes the sorting and (optionally) the
limiting applied when the backend generates statistics groups for Socket.IO
emission. Both sections can be extended without modifying TypeScript source
code.

\section{Go Client}

\subsection{Packet Capture}

The capture layer wraps the \texttt{gopacket/pcap} library. The
\texttt{capture.Start} function opens a live pcap handle on the specified
interface and applies an optional BPF filter. Code~\ref{lst:capture} shows
the core of the capture loop.

\begin{lstlisting}[language=Go, caption={Packet capture loop in the Go client.}, label={lst:capture}]
func Start(iface, bpf string, snapLen int, promisc bool) (
    *pcap.Handle, <-chan PacketEvent, error) {
  handle, err := pcap.OpenLive(iface, int32(snapLen), promisc,
                               pcap.BlockForever)
  if err != nil {
    return nil, nil, err
  }
  if bpf != "" {
    if err := handle.SetBPFFilter(bpf); err != nil {
      handle.Close()
      return nil, nil, fmt.Errorf("set BPF: %w", err)
    }
  }
  out := make(chan PacketEvent, 2048)
  src := gopacket.NewPacketSource(handle, handle.LinkType())
  go func() {
    defer close(out)
    for packet := range src.Packets() {
      // Extract IP layer, transport layer, and emit PacketEvent
      // ...
    }
  }()
  return handle, out, nil
}
\end{lstlisting}

If no BPF filter is provided by the operator, the client generates a default
filter string based on the capture \texttt{direction} flag (\texttt{out},
\texttt{in}, or \texttt{both}) and the local IPv4 addresses of the selected
interface. This default filter reliably excludes loopback and intra-host
traffic.

\subsection{Flow Aggregation and Deduplication}

The \texttt{Aggregator} type in \texttt{internal/flow/flow.go} maintains an
in-memory map keyed by a \texttt{Key} struct containing the five-tuple
(\texttt{SrcIP}, \texttt{DstIP}, \texttt{SrcPort}, \texttt{DstPort},
\texttt{Protocol}). When the \texttt{--dedupe ip} mode is selected, the key
collapses the port fields to zero so that multiple connections between the same
host pair count as a single flow entry.

Each map entry records first-seen and last-seen timestamps, cumulative byte
and packet counters split by direction (\texttt{bytesIn}, \texttt{bytesOut},
\texttt{packetsIn}, \texttt{packetsOut}), and a \texttt{dirty} flag indicating
uncommitted changes. Entries are marked inactive and evicted once they have
been idle for longer than the configured \texttt{--idle-ttl} duration.

At each flush interval, \texttt{ExportBatch} collects all dirty, non-pending
entries, serializes them into a \texttt{backend.ConnectionBatch} struct, and
sends an HTTP POST to \texttt{/api/connections}. The \texttt{pending} flag
prevents a flow from appearing in two consecutive batches before the first
request has completed.

\subsection{Configuration}

The client provides both command-line flags and environment variable equivalents
for every setting. Table~\ref{tab:client-flags} summarises the most significant
options.

\begin{table}[H]
  \centering
  \caption{Key configuration flags for the Go client.}
  \label{tab:client-flags}
  \begin{tabularx}{\textwidth}{llX}
    \toprule
    \textbf{Flag} & \textbf{Env var} & \textbf{Description} \\
    \midrule
    \texttt{--iface} & \texttt{BYTEROUTE\_IFACE} &
      Network interface to capture on (required). \\
    \texttt{--backend} & \texttt{BYTEROUTE\_BACKEND\_URL} &
      Backend base URL for ingestion requests. \\
    \texttt{--auth-token} & \texttt{BYTEROUTE\_AUTH\_TOKEN} &
      Bearer token used to authenticate ingestion requests. \\
    \texttt{--flush} & & Duration between batch POST requests (e.g., \texttt{5s}). \\
    \texttt{--direction} & & Capture direction: \texttt{out} (default), \texttt{in}, or \texttt{both}. \\
    \texttt{--bpf} & \texttt{BYTEROUTE\_BPF} &
      Custom BPF filter expression; overrides the generated default. \\
    \texttt{--dedupe} & & Deduplication mode: \texttt{flow} (five-tuple) or \texttt{ip} (IP pair). \\
    \texttt{--reporter-ip} & \texttt{BYTEROUTE\_REPORTER\_IP} &
      Public IP of the sensor host, enabling backend geo-location of private
      source addresses. \\
    \texttt{--idle-ttl} & & Maximum idle time before a flow entry is evicted. \\
    \texttt{--max-batch-conns} & & Maximum number of records per POST request. \\
    \bottomrule
  \end{tabularx}
\end{table}

\section{Dashboard}

The dashboard is a Vue.js 3 Single Page Application built with Vite. It
communicates with the backend through:

\begin{itemize}
  \item \textbf{REST API calls}: handled by composables and typed API client
    functions under \texttt{src/api/}, using the Fetch API or Axios.
  \item \textbf{Socket.IO}: a persistent connection managed by a dedicated
    service, receiving \texttt{trafficFlows} and \texttt{statisticsUpdate}
    events and propagating them to reactive Pinia stores.
\end{itemize}

Application state is managed with Pinia. Distinct stores handle
authentication state (user identity, token persistence), tenant selection,
connection records, and statistics data. Vue Router provides client-side
navigation between the map view, the connection list, and the statistics panels.

In production, Nginx serves the pre-built static bundle and proxies
\texttt{/api}, \texttt{/auth}, and \texttt{/socket.io} to the backend
container. The \texttt{VITE\_SOCKET\_URL} environment variable controls the
Socket.IO connection target: if unset, the client connects to the same origin,
which is correct in both the Vite dev-proxy scenario and the Docker Compose
scenario behind Traefik.

\section{Shared Package}

The \texttt{packages/shared} package exports:

\begin{itemize}
  \item TypeScript type definitions for \texttt{Connection}, \texttt{TrafficFlow},
    and related value objects, ensuring compile-time consistency across the
    backend and dashboard;
  \item Mongoose schema definitions and model constructors for
    \texttt{Connection}, \texttt{User}, and \texttt{Tenant}, allowing the
    backend to import and register them without duplicating schema code.
\end{itemize}

Because the shared package is linked as a local workspace dependency (declared
in both \texttt{apps/backend/package.json} and \texttt{apps/dashboard/package.json}),
any type change is immediately visible to all consumers without a publish step.
The \texttt{packages/typescript} package provides a shared
\texttt{tsconfig.base.json} used by all TypeScript projects in the monorepo.

% ============================================================
\chapter{DevOps and Quality Assurance}
\label{chap:devops}
% ============================================================

\section{DVCS Workflow}

ByteRoute uses \textbf{GitHub Flow}: development happens in short-lived feature
branches derived directly from \texttt{main}. No long-running integration or
release branches are maintained. Each change is merged through a Pull Request
(PR), which must pass all automated checks before merging. Direct pushes to
\texttt{main} are blocked at the repository level.

This strategy was chosen for its simplicity and its suitability for a small
team: it minimizes merge debt, keeps the production branch always deployable,
and integrates naturally with the automated release pipeline triggered on every
push to \texttt{main}.

\section{Conventional Commits and Semantic Release}

Every commit message must conform to the \textbf{Conventional Commits}
specification (\texttt{<type>[(scope)][!]: <description>}). Commitlint enforces
this format via a Husky \texttt{commit-msg} Git hook, ensuring that malformed
messages are rejected before reaching the remote.

Semantic Release analyses the commit history after each merge to \texttt{main}
and determines the next version number automatically:

\begin{itemize}
  \item \texttt{fix} commits trigger a patch version increment;
  \item \texttt{feat} commits trigger a minor version increment;
  \item commits containing a breaking change footer trigger a major version
    increment.
\end{itemize}

The release workflow then generates an updated \texttt{CHANGELOG.md}, creates
a GitHub Release with the computed version tag, and (via
\texttt{@semantic-release/git}) commits the version bump back to the repository.
A \texttt{sync-workspace-versions.mjs} script propagates the new version to
all \texttt{package.json} files in the monorepo.

\section{Quality Assurance Tools}

\subsection{Husky and Lint-staged}

Husky manages two pre-commit hooks:

\begin{itemize}
  \item \texttt{commit-msg}: runs Commitlint and rejects commits that do not
    follow the Conventional Commits format.
  \item \texttt{pre-commit}: runs Lint-staged, which applies ESLint and
    Prettier only to the files staged for the current commit, keeping the
    feedback cycle fast without reformatting the entire codebase.
\end{itemize}

\subsection{ESLint and Prettier}

ESLint configuration files (\texttt{eslint.config.ts} in the backend,
\texttt{eslint.config.js} in the dashboard) enforce code quality rules
consistent with the TypeScript style guide. Prettier is configured at the
workspace root and ensures a uniform formatting style across all
JavaScript/TypeScript source files.

\subsection{Renovate}

A Renovate configuration (\texttt{renovate.json}) scans the repository for
outdated dependencies and opens automated PRs to update them. This keeps the
dependency graph current without requiring manual tracking of upstream releases.

\section{Continuous Integration and Delivery}

The CI/CD pipeline is implemented as GitHub Actions workflows. Two main
workflows are defined:

\textbf{Build and Test} (\texttt{build-and-test.yml}): triggered on every PR
and push to \texttt{main}. It:
\begin{enumerate}
  \item Installs dependencies with \texttt{pnpm install --frozen-lockfile};
  \item Runs TypeScript type checks (\texttt{tsc --noEmit}) for the backend and
    the dashboard;
  \item Executes backend unit tests and end-to-end tests with Vitest, generating
    Cobertura XML coverage reports;
  \item Builds and runs Go client tests, generating a \texttt{coverage.out} file
    and converting it to Cobertura XML;
  \item Merges backend unit and e2e coverage reports using a custom merge script;
  \item Publishes coverage artifacts (\texttt{coverage-reports},
    \texttt{coverage-go}) and displays a Markdown summary in the GitHub Actions
    job summary via the \texttt{irongut/CodeCoverageSummary} action;
  \item Uploads coverage data to Codecov for badge generation.
\end{enumerate}

\textbf{Release} (\texttt{release.yml}): triggered on push to \texttt{main}
after all checks pass. It:
\begin{enumerate}
  \item Runs Semantic Release to calculate the next version and create the
    GitHub Release;
  \item Conditionally builds Docker images for the backend, dashboard, and Go
    client (only when a new release is published, to avoid creating redundant
    image tags for non-version commits such as documentation updates);
  \item Tags images with both the semantic version (e.g., \texttt{1.4.2}) and
    the \texttt{latest} label, then pushes them to the GitHub Container Registry
    (\texttt{ghcr.io/ste-script}).
\end{enumerate}

% ============================================================
\chapter{Validation and Results}
\label{chap:validation}
% ============================================================

\section{Testing Strategy}

The test suite covers the backend at two levels:

\begin{itemize}
  \item \textbf{Unit tests}: located under \texttt{apps/backend/test/}, they
    cover individual service functions (ingestion logic, GeoIP enrichment,
    authentication, tenant resolution) in isolation, using mock objects where
    infrastructure dependencies would otherwise be required.
  \item \textbf{End-to-end tests}: also located under
    \texttt{apps/backend/test/}, they spin up an in-memory MongoDB instance
    (via \texttt{mongodb-memory-server}) and verify the full request life cycle
    from the HTTP router through to database state, including authentication,
    connection ingestion, and tenant isolation.
\end{itemize}

The test runner is \textbf{Vitest}, configured in \texttt{vitest.config.ts}.
The end-to-end configuration is kept separate (\texttt{vitest.config.ts} with
an \texttt{e2e} project entry) to allow the two test types to run independently
or together.

The Go client uses the standard \texttt{go test} toolchain. Tests exist for the
BPF filter generation logic (\texttt{capture/bpf\_test.go}) and for the flow
aggregation and deduplication logic (\texttt{flow/flow\_test.go}).

The dashboard uses Vitest with \texttt{@vue/test-utils} for component and
composable tests, located under \texttt{apps/dashboard/src/test/}.

\section{Coverage}

Coverage is collected at three levels:

\begin{enumerate}
  \item \textbf{Backend unit} and \textbf{end-to-end} coverage is gathered by
    Vitest using V8 instrumentation and exported as Cobertura XML. The two
    reports are merged by the \texttt{scripts/merge-coverage.mjs} script into a
    single consolidated file, which is uploaded to Codecov and archived as a CI
    artifact alongside the HTML report generated by
    \texttt{scripts/coverage-e2e.mjs}.
  \item \textbf{Go client} coverage is produced by \texttt{go test --coverprofile}
    and converted to Cobertura XML by \texttt{gocover-cobertura}. The Cobertura
    file and the raw \texttt{coverage.out} are published as the \texttt{coverage-go}
    CI artifact.
  \item Both Cobertura reports are consumed by the
    \texttt{irongut/CodeCoverageSummary} GitHub Action, which posts a formatted
    Markdown table to the job summary, making coverage data immediately visible
    in the Actions interface without downloading artifacts.
\end{enumerate}

The overall coverage level is tracked via a Codecov badge displayed in the
repository \texttt{README.md}, providing a persistent, public quality signal.

\section{Observable System Behavior}

End-to-end correctness is also validated qualitatively through integration
testing with the running Docker Compose stack:

\begin{itemize}
  \item A \texttt{scripts/seed.ts} script wipes the database and inserts a
    baseline user and tenant, enabling a reproducible starting state for manual
    and automated acceptance tests.
  \item The \texttt{scripts/ensure-connection-indexes.mjs} script verifies that
    MongoDB compound indexes (\texttt{tenantId} + \texttt{id},
    \texttt{tenantId} + \texttt{startTime}) are present, preventing silent
    performance degradation in larger deployments.
  \item The backend exposes a \texttt{GET /health} endpoint, which returns an
    HTTP 200 response when the service is ready to accept traffic. This endpoint
    is used by Docker Compose health checks and CI smoke tests.
\end{itemize}

% ============================================================
\chapter{Conclusion}
\label{chap:conclusion}
% ============================================================

ByteRoute demonstrates the viability of a full-stack, self-hosted network
traffic monitoring platform built on modern open-source technologies. The
combination of a lightweight Go sensor, a Node.js enrichment and broadcasting
backend, and a Vue.js 3 dashboard provides a complete real-time visibility
solution that can be deployed with a single \texttt{docker compose up} command.

From an engineering standpoint, the project exhibits several qualities that are
indicative of sound software design practice:

\begin{itemize}
  \item \textbf{Separation of concerns}: the monorepo layout and the shared
    typed contracts ensure that the three main components are independently
    developed and tested while remaining interoperable.
  \item \textbf{Extensibility}: the domain DSL externalizes behavioral rules,
    allowing operators to adjust ingestion filtering and analytics parameters
    without rebuilding the application.
  \item \textbf{Reproducibility}: the Docker Compose setup, the frozen lockfile,
    and the seed script together guarantee that any developer can obtain an
    identical running environment in a few minutes.
  \item \textbf{Automation}: Conventional Commits, Semantic Release, and GitHub
    Actions reduce manual overhead at every stage of the delivery pipeline, from
    commit message validation to Docker image publication.
\end{itemize}

Future work could extend the platform in several directions. Support for
additional capture backends (e.g., eBPF-based sensors) would reduce the
privilege requirements of the client. A time-series storage backend (such as
InfluxDB or TimescaleDB) would improve the scalability and queryability of
historical traffic data. An alerting subsystem, modeled on a push notification
architecture similar to that of AlmaSpot, could notify administrators when
traffic to unexpected ASNs or countries is detected. Finally, expanding the
dashboard with per-connection detail views and exportable reports would broaden
the platform's applicability to security auditing and compliance use cases.

\end{document}
